Stochastic blockmodeling of relational event dynamics

### 2012-06-01 docs/discussion.tex
Fix up some references.

\begin{figure}[t]
\centering
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/observed}
\caption{Observed counts}
\end{subfigure}
~
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/1}
\caption{Intercept estimates}
\end{subfigure}
~
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/2}
\caption{\texttt{ab-ba} estimates}
\end{subfigure}
~
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/3}
\caption{\texttt{ab-by} estimates}
\end{subfigure} \\
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/observed-zoom}
\caption{Observed counts}
\end{subfigure}
~
\begin{subfigure}[b]{0.22\textwidth}
\end{subfigure}
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/1-zoom}
\caption{Intercept estimates}
\end{subfigure}
~
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/2-zoom}
\caption{\texttt{ab-ba} estimates}
\end{subfigure}
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/3-zoom}
\caption{\texttt{ab-by} estimates}
\end{subfigure}
\caption{Comparing observed counts and parameter estimates.  Darker values are larger.  Estimates are rescaled posterior means $(\hat{\beta}_{p,z_i,z_j} - \hat{\mu}_p)/\hat{\sigma}_p$ for each dyad $(i,j)$.  Learned parameters suggest heterogeneity exists in both total activity (b) as well as dynamics, as seen in (c) and (d).  Figures (e-h) give a zoomed-in view of groups 1 and 3, showing that while the overall rate of (1,3) and (3,1) events is similar, the tendency for \texttt{ab-by} differs.}
\label{fig:parmats}
\end{figure}

>>>>

\begin{figure}[t]
\centering
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/observed}
\caption{Observed counts}
\end{subfigure}
~
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/1}
\caption{Intercept estimates}
\end{subfigure}
~
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/2}
\caption{\texttt{ab-ba} estimates}
\end{subfigure}
~
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/3}
\caption{\texttt{ab-by} estimates}
\end{subfigure} \\
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/observed-zoom}
\caption{Observed counts}
\end{subfigure}
~
\begin{subfigure}[b]{0.22\textwidth}
\end{subfigure}
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/1-zoom}
\caption{Intercept estimates}
\end{subfigure}
~
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/2-zoom}
\caption{\texttt{ab-ba} estimates}
\end{subfigure}
\begin{subfigure}[b]{0.22\textwidth}
\centering
\includegraphics[scale=.25]{../figs/eckmann-small/parmat/3-zoom}
\caption{\texttt{ab-by} estimates}
\end{subfigure}
\caption{Comparing observed counts and parameter estimates.  Darker values are larger.  Estimates are rescaled posterior means $(\hat{\beta}_{z_i,z_j,p} - \hat{\mu}_p)/\hat{\sigma}_p$ for each dyad $(i,j)$.  Learned parameters suggest heterogeneity exists in both total activity (b) as well as dynamics, as seen in (c) and (d).  Figures (e-h) give a zoomed-in view of groups 1 and 3, showing that while the overall rate of (1,3) and (3,1) events is similar, the tendency for \texttt{ab-by} transitions differs.}
\label{fig:parmats}
\end{figure}

### 2012-06-01 docs/experiments.tex
Edit discussion and reorganize a few things.
Small edits.
Fix up some references.

\begin{itemize}
%\item Kiel email: \cite{Ebel2002}
\item Classrooms: TODO describe datasets. Each have about 20 people and 400 events.  \cite{McFarland2003}
\item Eckmann email: dyadic emails.  2000 training events among 88 people and 1300 test events.  \cite{Eckmann}
\item Enron email: dyadic emails between months [TODO], 3000 training and 1000 test events among 141 actors \cite{Enron}
\item Irvine: dyadic interactions among 401 actors each having more than 30 events,  \cite{Opsahl}
\item Tweets from Twitter.com occurring between from May 11, 2009 to January 26, 2012 that contained the hashtag \texttt{\#rstats}.\footnote{Will be made publicly available.}  This hashtag is used to denote messages pertaining to the R statistical computing environment and sometimes statistical discussion more generally.  We collect dyadic events by selecting tweets beginning with the \texttt{@} symbol (called a \emph{mention}), and mark the first mentioned user as the recipient.
Of 28337 total tweets in this time period, 3926 were directed events among a total of 1079 users.
We use the subset of 487 users who participated in more than one event, using a training set of 2000 events and a test set of 1330 events.
\item MIT Reality Mining: phone calls among the 89 recipients between the dates , using of 2000 training events followed by 1000 test events. \cite{MITreality}
\end{itemize}

@@

\input{../figs/results-llk.tex}

\subsection{Baselines}

@@

\subsection{Results}

TODO: Quickly describe main take-aways from the results tables.

>>>>

\begin{itemize}
%\item Kiel email: \cite{Ebel2002}
\item Classrooms: TODO describe datasets. Each have about 20 people and 400 events.  \cite{McFarland2001}
\item Eckmann email: dyadic emails.  2000 training events among 88 people and 1300 test events.  \cite{Eckmann2004}
\item Enron email: dyadic emails between months [TODO], 3000 training and 1000 test events among 141 actors \cite{Klimt2004}
%\item Irvine: dyadic interactions among 401 actors each having more than 30 events,  \cite{Opsahl}
\item Tweets from Twitter.com occurring between from May 11, 2009 to January 26, 2012 that contained the hashtag \texttt{\#rstats}.\footnote{Will be made publicly available.}  This hashtag is used to denote messages pertaining to the R statistical computing environment and sometimes statistical discussion more generally.  We collect dyadic events by selecting tweets beginning with the \texttt{@} symbol (called a \emph{mention}), and mark the first mentioned user as the recipient.
Of 28337 total tweets in this time period, 3926 were directed events among a total of 1079 users.
We use the subset of 487 users who participated in more than one event, using a training set of 2000 events and a test set of 1330 events.
\item MIT Reality Mining: phone calls among the 89 recipients between the dates , using of 2000 training events followed by 1000 test events. \cite{Eagle2009}
\end{itemize}

@@

\input{../figs/results-llk2.tex}

%\subsection{Baselines}

@@

### 2012-06-01 docs/intro.tex
Fix up some references.

In the cross-sectional case, \emph{stochastic blockmodels} \cite{Nowicki2001, Kemp, Ishiguro2010, Rodriguez} have been proposed as a family of approaches that capture behavioral similarity by identifying subsets of nodes with similar patterns of ties to those in other sets.  While it is natural to apply these ideas directly to relational event data via blockmodeling of the time-marginalized rates of interaction among dyads (effectively treating the event structure as a valued graph), there are limits to what this approach can detect.  Consider, e.g., Figure~\ref{fig:introexample}.  In the top left panel, we depict the time-marginalized frequencies of simulated interactions between members of two groups (A and B), with darker cells indicating higher interaction frequencies.  As can be seen, a clear role structure is present, with members of each subgroup interacting at higher rates with co-members than out-group members; such a structure is easily detectable via conventional blockmodeling techniques.  By contrast, consider the interaction patterns shown in the top right panel.  Here, there is no systematic difference in marginal rates between the two groups, rendering them invisible under a standard blockmodeling approach.

@@

\begin{SCfigure}
\centering
\includegraphics[scale=.45]{../figs/introexample/all}
\caption{Illustration of differentiation by rate of interaction versus by dynamic behavior.  The left column shows an example where within-block communication is large, and there is a higher tendency for reciprocity within a block.  In the situation where the two groups are not differentiated by rate, as in the right column, a standard blockmodel is unable to distinguish between groups A and B.  The proposed method can, however, learn such groups by employing more flexible definitions of shared dynamics.}
\label{fig:example}
\end{SCfigure}
% {\color{red} (Suggest renaming "Dynamics parameters" to "Behavioral parameters" or something.  "Dynamics parameters" is not gramatically sound.)

>>>>

In the cross-sectional case, \emph{stochastic blockmodels} \cite{Nowicki2001, Kemp, Ishiguro2010, Rodriguez2011} have been proposed as a family of approaches that capture behavioral similarity by identifying subsets of nodes with similar patterns of ties to those in other sets.  While it is natural to apply these ideas directly to relational event data via blockmodeling of the time-marginalized rates of interaction among dyads (effectively treating the event structure as a valued graph), there are limits to what this approach can detect.  Consider, e.g., Figure~\ref{fig:introexample}.  In the top left panel, we depict the time-marginalized frequencies of simulated interactions between members of two groups (A and B), with darker cells indicating higher interaction frequencies.  As can be seen, a clear role structure is present, with members of each subgroup interacting at higher rates with co-members than out-group members; such a structure is easily detectable via conventional blockmodeling techniques.  By contrast, consider the interaction patterns shown in the top right panel.  Here, there is no systematic difference in marginal rates between the two groups, rendering them invisible under a standard blockmodeling approach.

@@

\begin{SCfigure}
\centering
\includegraphics[scale=.45]{../figs/introexample/all}
\caption{Illustration of differentiation by rate of interaction versus by dynamic behavior.  The left column shows an example where within-block communication is large, and there is a higher tendency for reciprocity within a block.  In the situation where the two groups are not differentiated by rate, as in the right column, a standard blockmodel is unable to distinguish between groups A and B.  The proposed method can, however, learn such groups by employing more flexible definitions of shared dynamics.}
\label{fig:introexample}
\end{SCfigure}
% {\color{red} (Suggest renaming "Dynamics parameters" to "Behavioral parameters" or something.  "Dynamics parameters" is not gramatically sound.)
